### 2014-02-23 presentation-plan/presentation-plan.tex
First start for the summary of Incremental Aggregation Plan

\documentclass[runningheads,a4paper]{llncs}

\usepackage{amssymb}
\setcounter{tocdepth}{3}
\usepackage{graphicx}

\usepackage{url}
\urldef{\mailsa}\path|{i.c.t.m.speek, a.c.stolwijk}@student.tudelft.nl|
\newcommand{\keywords}[1]{\par\addvspace\baselineskip%
\noindent\keywordname\enspace\ignorespaces#1}

\begin{document}

\mainmatter% start of an individual contribution

% first the title is needed
\title{Presentation Plan\\
FF-Replan: \& RFF\@: Exploiting classical AI planning for uncertain and probabilistic domains}

% a short form should be given in case it is too long for the running head
\titlerunning{Lecture Notes in Computer Science: Authors' Instructions}

\author{I.C.T.M Speek, A.C. Stolwijk}

%
\authorrunning{Presentation Plan Seminar Algorithms FF-replan \& RFF}
% (feature abused for this document to repeat the title also on left hand pages)

% the affiliations are given next; don't give your e-mail address
% unless you accept that it will be published
\institute{Seminar Algorithms, Embedded Software,
Msc Embedded Systems,\\
Delft University of Technology\\
\mailsa\\
}

\toctitle{Lecture Notes in Computer Science}
\tocauthor{Authors' Instructions}
\maketitle

\section{Summary of given papers}

\subsection{FF-Replan: A Baseline for Probabilistic Planning}

At the first international probabilistic planning competition IPPC-04, most
entries were Markov Decision Processes. The winner, FF-Replan, however was
based on deterministic planning techniques.

By relying on search, FF-Replan could function with enormous speed and efficiency.

Until the paper there have been two IPPC conferences.

\subsection{Incremental Plan Aggregation for Generating Policies in MDPs}

Markov Decision Processes (MDPs) is the best known formalism for planning under
uncertainty. It can however be computationally intensive if large parts of the
policies are tried to optimize at once. Most recent MDP planning techniques are
using the following steps as a basis approach:

\begin{enumerate}
	\item Take MDP problem
	\item ``Determize''  it into a classical planning problem
	\item Generate sequence of actions that represent a possible exectution
			path
	\item MDP planner executes plan:
		\begin{itemize}
			\item If goal is reached, reports success
			\item Otherwise algorithm generates a new plan from failed state of the world
		\end{itemize}
\end{enumerate}

\section{Questions that arose}

\section{Criticism on the contribution of the paper}

\section{Relevant literature on the subject}

\section{Presentation setup}

\end{document}