### 2014-02-23 presentation-plan/presentation-plan.tex
I think the summary includes everything now.

\documentclass[runningheads,a4paper]{llncs}

\usepackage{amssymb}
\setcounter{tocdepth}{3}
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage{fixltx2e}

\usepackage{url}
\urldef{\mailsa}\path|{i.c.t.m.speek, a.c.stolwijk}@student.tudelft.nl|
\newcommand{\keywords}[1]{\par\addvspace\baselineskip%
\noindent\keywordname\enspace\ignorespaces#1}

\begin{document}

\mainmatter% start of an individual contribution

% first the title is needed
\title{Presentation Plan\\
FF-Replan: \& RFF\@: Exploiting classical AI planning for uncertain and probabilistic domains}

% a short form should be given in case it is too long for the running head
\titlerunning{Presentation Plan}

\author{I.C.T.M Speek, A.C. Stolwijk}

%
\authorrunning{Presentation Plan Seminar Algorithms FF-replan \& RFF}
% (feature abused for this document to repeat the title also on left hand pages)

% the affiliations are given next; don't give your e-mail address
% unless you accept that it will be published
\institute{Seminar Algorithms, Embedded Software,
Msc Embedded Systems,\\
Delft University of Technology\\
\mailsa\\
}

%\toctitle{Lecture Notes in Computer Science}
%\tocauthor{Authors' Instructions}

\maketitle

\section{Summary of given papers}

\subsection{FF-Replan: A Baseline for Probabilistic Planning}

At the first international probabilistic planning competition IPPC-04, most
entries were Markov Decision Processes. The winner, FF-Replan, however was
based on deterministic planning techniques.

By relying on search, FF-Replan could function with enormous speed and efficiency.

Until the paper there have been two IPPC conferences.

\subsection{Incremental Plan Aggregation for Generating Policies in MDPs}

Markov Decision Processes (MDPs) is the best known formalism for planning under
uncertainty. It can however be computationally intensive if large parts of the
policies are tried to optimize at once. Most recent MDP planning techniques are
using the following steps as a basis approach:

\begin{enumerate}
	\item Take MDP problem
	\item ``Determinize''  it into a classical planning problem
	\item Generate sequence of actions that represent a possible execution
		path
	\item MDP planner executes plan:
		\begin{itemize}
			\item If goal is reached, reports success
			\item Otherwise algorithm generates a new plan from failed state of
				the world
		\end{itemize}
\end{enumerate}

There are some issues with this approach:

\begin{itemize}
	\item It stops after the first computed path to a goal which is likely to fail
	\item May result in generating the same path over and over again
	\item Only generates an execution path, not a policy
\end{itemize}

The authors identified four aspects of policy generation for MDP planning
problems using a classical planning algorithm:

\begin{enumerate}
	\item How to determinize the MDP planning problem
	\item which plan (initial state, goal) to ask for
	\item How to incrementally aggregate the plans by the classical planner
		into a solution policy
	\item How to combine this coherently
\end{enumerate}

The paper presents the Robust FF (\texttt{RFF}) planner, which should solve the
issues. \texttt{RFF} first generates an initial plan from the initial state to
some goal using \texttt{FF} on a determination of the MDP\@. There are several
strategies to do this. \textsc{MostProbableOutcome} (MPO) only uses the action
from a state with the highest probability. \textsc{AllOutcomes} (AO) uses all
actions from a state. Then for each state in the plan, the successors of that
state are added too. This is expanding the execution structure. For the new
state the probability to reach this state should be higher than a certain
threshold and if \texttt{FF} does not return a failure for this state and a
certain goal. The goals can be chosen in three different ways, using
\textsc{ProblemGoal}, which uses the goals of the MDP, \textsc{RandomGoal},
which randomly picks some states from the policy graph and \textsc{BestGoal}
which picks the best states from the policy graph. The calculation of the
probability to a state from the initial state can be costly. To optimize this
in \texttt{RFF} a Monte-Carlo (MC) simulation is used. \texttt{RFF} terminates
if either the global probability is lower than the threshold, or there are no
new terminal states to be added.

Three theorems, one lemma and one corollary about MDP problems using
\texttt{RFF} are given:

\begin{description}
	\item[Lemma 1] For every MDP planning problem RFF terminates in
		finite time
	\item[Theorem 1] (Soundness of \texttt{RFF\textsubscript{MPO}}). For every
		MDP planning problem, every solution that
		\texttt{RFF\textsubscript{MPO}} finds is correct
	\item[Theorem 2] (Completeness of \texttt{RFF\textsubscript{AO}}). For
		every MDP planning problem, if a solution exists, it is found by
		\texttt{RFF\textsubscript{AO}} or returns a failure
	\item[Corollary 1] (Soundness of \texttt{RFF\textsubscript{AO}}). For every
		MDP planning problem, every solution that
		\texttt{RFF\textsubscript{AO}} finds is correct
	\item[Theorem 3] If there are no unsolvable states in the MDP, then the
		probability of success of any solution found by \texttt{RFF} is
		higher than $1 - \rho$ where $\rho$ is the threshold probability
\end{description}

For the IPC-08 the authors have implemented \texttt{RFF} in C++ to perform
experiments. It performed experiments on the ``fully-observable probabilistic
(FOP) planning track''. At IPC-08 it performed better than other competitors.
The experiments confirmed Theorem 3, a higher threshold mean that \texttt{RFF}
will generate policies that are more prone to replanning.

In the experiments different Goal-Selection strategies were tested.
\textsc{ProblemGoal} (PG) performed the worst. \textsc{RandomGoal} performed
better and \textsc{BestGoal} generally produced higher quality goals, but also
took more time.

The different ways of determinizing MDPs was also tested. It turned out that
\textsc{AllOutcomes} solve much larger problems. However
\textsc{MostProbableOutcome} generally solve more problems in less time.

\section{Questions that arose}

\section{Criticism on the contribution of the paper}

\section{Relevant literature on the subject}

\section{Presentation setup}

\end{document}