### 2014-03-11 paper-plan/paper-plan.tex
added setup for paper

\documentclass[runningheads,a4paper]{llncs}

\usepackage{amssymb}
\setcounter{tocdepth}{3}
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage{fixltx2e}

\usepackage{url}
\urldef{\mailsa}\path|{i.c.t.m.speek, a.c.stolwijk}@student.tudelft.nl|
\newcommand{\keywords}[1]{\par\addvspace\baselineskip%
\noindent\keywordname\enspace\ignorespaces#1}

\begin{document}

\mainmatter% start of an individual contribution

% first the title is needed
\title{Paper Plan\\
FF-Replan: \& RFF\@: Exploiting classical AI planning for uncertain and probabilistic domains}

% a short form should be given in case it is too long for the running head
\titlerunning{Presentation Plan}

\author{I.C.T.M Speek, A.C. Stolwijk}

%
\authorrunning{Paper Plan Seminar Algorithms FF-replan \& RFF}
% (feature abused for this document to repeat the title also on left hand pages)

% the affiliations are given next; don't give your e-mail address
% unless you accept that it will be published
\institute{Seminar Algorithms, Embedded Software,
Msc Embedded Systems,\\
Delft University of Technology\\
\mailsa\\
}

%\toctitle{Lecture Notes in Computer Science}
%\tocauthor{Authors' Instructions}

\maketitle

%--------------------------------------------------------------------

\section{Introduction}

\begin{itemize}
	\item Introduce planning problem
	\item Introduce planning for uncertainty
	\item Introduce the problem definition and motivation
\end{itemize}

There is some domain in which requires probabilistic planning. As FF-Replan and
RFF are about using deterministic planners in a probabilistic domain, we like
to see what's new and if we can find some holes that are still open.

\section{Related Work}

%---------------------------------------------------------------------

\section{Related works}

\begin{itemize}
	\item Introduce the previously read papers
	\item Introduce the current state of the art papers that use deterministic approaches for probablistic domains deadling with uncertainty
\end{itemize}

\subsubsection{Probabilistic Planning in the Graphplan Framework (1999)~\cite{Blum99probabilisticplanning}}

Graphplan is a successful planning algorithm for classical STRIPS domains. This
paper explores the extend it can be used in \emph{probabilistic} domains. The
paper discusses two variations of Graphplan: PGraphplan and TGraphplan.
PGraphplan produces an optimal plan, while TGraphplan produces a sub-optimal
plan but it has increased speed. By comparing the speed and quality of the
two planners the authors are able to estimate how far they are from the ideal.

\subsubsection{Compiling Conformant Probabilistic Planning Problems into Classical Planning (2013)~\cite{taig2013conformant}}

%---------------------------------------------------------------------

\subsection{How Much Does a Household Robot Need To Know In Order To Tidy Up? (2013)~\cite{nebel2013much}}

For household robot planning, it appears easy, but actually involves
uncertainty. For tidying things up objects might not be there, or sensing
operations might tell things wrong. This paper looks into conditions for
classical planning in a replanning loop in order to solve nondeterministic
partially observable open domains This paper looks into conditions for
classical planning in a replanning loop in order to solve nondeterministic
partially observable open domains.

\section{Problem formulation}
\begin{itemize}
	\item Summarize shortcomings of the previous solutions
	\item Describe the planning problem
	\item Argument the choice for 2 current papers to analyze
	\item Define how we would like to test the planning algorithms in a domain by our choice
\end{itemize}

From the related literature we identify some problems.

%---------------------------------------------------------------------

\section{Domain description}

\begin{itemize}
	\item Describe domains from IPPC
	\item Explain shortcomings
	\item Evaluate a valuable real-world domain in which we can test 2 papers
\end{itemize}

%---------------------------------------------------------------------

\section{Theoretical approach}

\begin{itemize}
	\item Describe how we are using the created domain
	\item show pseudo code for the 2 papers
	\item Introduce the improvements (or combination) of the approaches of the papers fitted for the domain
\end{itemize}

%---------------------------------------------------------------------

\section{Empirical evalution}

\begin{itemize}
	\item Show results for the experiment
	\item Critically analyze the results
	\item Compare with results from the original papers
\end{itemize}

%---------------------------------------------------------------------

\section{Conclusion}

\begin{itemize}
	\item Conclude on the succes of the 2 approaches
	\item Conclude the work done in the paper
	\item Hint to some future work developments
\end{itemize}

\begin{itemize}
	\item \cite{Hoffmann01theff}
	\item \cite{FFReplan}
	\item \cite{teichteil2010incremental}
	\item \cite{yoon2008probabilistic}
	\item \cite{teichteil2012fast}
	\item \cite{bonet2011planning}
	\item \cite{taig2013translation}
	\item \cite{martin2013progressive}
	\item \cite{taig2013conformant}
	\item \cite{2013workshop}
\end{itemize}

\bibliographystyle{plain}
\bibliography{paper-plan}

\end{document}