### 2014-04-09 FinalPaper/paper.tex

\documentclass[runningheads,a4paper]{llncs}

\usepackage{amssymb}
\setcounter{tocdepth}{3}
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage{fixltx2e}
\usepackage{cite}

\usepackage{xcolor}
\newcommand\myworries[1]{\textcolor{red}{#1}}
\newcommand\todo[1]{\textcolor{blue}{#1}}

\usepackage{url}
\urldef{\mailsa}\path|{i.c.t.m.speek, a.c.stolwijk}@student.tudelft.nl|
\newcommand{\keywords}[1]{\par\addvspace\baselineskip%
\noindent\keywordname\enspace\ignorespaces#1}

\begin{document}

\mainmatter% start of an individual contribution

% first the title is needed
\title{Realistically Interesting Domains for Probabilistic Planners}

% a short form should be given in case it is too long for the running head
\titlerunning{Real World Domains}

\author{I.C.T.M Speek, A.C. Stolwijk}

%
\authorrunning{Validating Planners in a Real World Domain}
% (feature abused for this document to repeat the title also on left hand pages)

% the affiliations are given next; don't give your e-mail address
% unless you accept that it will be published
\institute{Seminar Algorithms, Embedded Software,
MSc Embedded Systems,\\
Delft University of Technology\\
\mailsa\\
}

%\toctitle{Lecture Notes in Computer Science}
%\tocauthor{Authors' Instructions}

\maketitle

%--------------------------------------------------------------------

\begin{abstract}
	In contrast to current demand for probabilistically interesting domains in the International Probabilistic Planning Competitions, this papier argues for using real-world appropriate domains. By relaxing the definition of a dead-end, we argue that for many
real life domains a simpler planning algorithm can be sufficient and even better than sophisticated probabilistic planners.

In this paper we evaluate the strength of FF-Replan, winner of IPPC-04 and unenrolled winner of IPPC-06, a replanner that uses the classical planner FF on a determinized probabilistic problem against a state of the art probabilistic planner \textsc{Gourmand}, the successor of the Labelled Real-Time Dynamic Probabilistic planner \textsc{Glutton}, runner up of the IPPC-11. By translating the Navigation and Traffic domain from IPPC-11 to a matching real-world equivalent and evaluating these using the proposed probabilistically interesting baseline test as proposed by Little et al\. we evaluate whether FF-Replan is in fact capable enough to solve enough of these so called classical planning problems in disguise.
\end{abstract}

%--------------------------------------------------------------------

\section{Introduction}
The introduction of the International Planning Competition (IPC) in 1998 furthered the field of automated planning and scheduling. Since then several state-of-the-art planners such as the FF-planner \cite{Hoffmann01theff} have been introduced optimized for the initial classical planning problems. Surprisingly well-performing planners have caused the field to readjust the then supposedly challenging domains to provide a challenge for a wide variety of planners. By doing so, the definition of an interesting planning problem has been modified.

The entrance of FF-Replan \cite{FFReplan}, a replanner based on determinization techniques and the classical FF planner, surprisingly outperformed other participating MDP solvers in the International Probabilistic Planning Competition (IPPC) '04. As a result, Little et al. have argued for \emph{probabilistically interesting} domains \cite{little2007probvsreplan} and finite horizon MDPs were employed in IPPC-11. These approximations of infinite planning problems however strongly resemble optimization problems rather than the goal oriented initial planning domains.  This shift in the current interesting domains caused the state-of-the-art planners to cope with the large branching factors that are introduced using finite horizon MDPs.

This paper presents a new domain involving Unmanned Aerial Vehicles (UAVs) cooperating to deliver data packages from one UAV to the base station. To match this problem with an appropriate planner, this paper presents an overview of other interesting domains and probabilistic planners that were involved in the redefinition of an interesting planning problem. By evaluating the domains with the help of the probabilistic interestingness baseline test introduced by Little et al. \cite{little2007probvsreplan} a recommendation will be made for a suitable planner to match the UAV problem.

Section \ref{sec:problem} introduced the problem formulation this paper addresses while Section \ref{sec:domaind} introduces the details of proposed domain. Section \ref{sec:background} introduced the appropriate prior knowledge and Section \ref{sec:domains} and \ref{sec:planners} the interesting domains and planners respectively. Section \ref{sec:experiment} briefly elaborates on a possible experiment set-up. Finally Section \ref{sec:conclusion} concludes on the evaluation of the domain and an appropriate planner.

%---------------------------------------------------------------------

\section{Problem formulation}
\label{sec:problem}

To evaluate FF-Replan and \textsc{Gourmand} in a real world domain a finite horizon MDP is used to represent the realistic \emph{natural dynamics} such as external forces or forces of nature. These dynamics introduce large branching factors emphasizing the need for a planner to be scalable. To evaluate the model, the planner should also be sensitive to an expected reward maximization criterion.
Current finite horizon MDP problems as introduced in IPPC-11 are simplified and therefor do not account for the rich set of state-actions as would be available in a real-world planning problem. We will reason about the most interesting domains for our evaluation and adjust them to match their real-world equivalent. Using the baseline test introduced by Little et al.\@ \cite{little2007probvsreplan}, we can evaluate the probabilistically interestingness of these domains and evaluate the performance of FF-Replan vs.\@ a state of the art probabilistic planner: \textsc{Gourmand}.

\section{Domain Description}
\label{sec:domaind}

Unmanned Aerial Vehicles (UAVs), or commonly called \emph{Drones} are aircrafts
without a pilot and often autonomously controlled. Drones are widely used for
military as well as for civil uses, including inspecting pipelines, counting
wildlife, agriculture, filming or delivering supplies to inaccessible areas.
As many uses cases there are, there are many forms of UAVs, bigger fixed wing
aircrafts, or small helicopter or quad copter like aircrafts that can hover at
a certain position in the air.

We are particularly interested in monitoring, for example crops or livestock,
and surveillance UAVs that work together to monitor a large area. Each UAV
monitors a part of the total area, and if it has monitored something, it has
to communicate this to a base station. This can't be done directly due to the
large distance, instead it has to go through neighboring UAVs.

Communicating the message from the sensing UAV to the base station can be
expressed as a probabilistic planning problem. Initially the sensing UAV
has the package that has to be sent, and the goal is to bring this package to
the base station. Each time the package is passed to the neighboring UAV,
there is a possibility the package is not received correctly. The bigger the
distance, the higher the probability the delivery failed.

\todo{TODO maybe UAVs can be on or of too, so it's forced to use a further UAV.}
\todo{TODO what's also interesting is that Amazon.com uses drones for package deliveries}

%---------------------------------------------------------------------

\section{Background}
\label{sec:background}

\subsection{Planning problems}
\subsubsection{Classical Planning problems. } Planning is the process of computing several steps of a problem-solving procedure before executing any of them. In classical planning problems the environment is assumed to \emph{deterministic, observable and state}. Classical planning problems consists out of states, actions, conditions and a goal state and are solved using search and can be aided using heuristics. The planner tries to find a sequence of actions such that a goal or maximum reward is achieved.

\subsubsection{Probabilistic Planning problems}
Real-World applications are far from deterministic, induced by natural dynamics and chance. To accommodate for Real-World domains probabilistic planning problems were introduced.  These systems can be described using \emph{Markov Decision Processes} (MDPs) and can account for observable uncertainty. Probabilistic planning problems can be solved by evaluating iterations and creating policies to reach a goal state, either online or offline.

\subsection{Problem domains}
\subsubsection{Goal-oriented MDPs. }The problems in IPPC-04 were specified using the Planning Domain Description Language (PPDLL). PPDDL specifies a probabistic planning problem as a set of probabilistic action sets and an initial and goal state description in which every action has a certain probability assigned to it \cite{FFReplan}. These MDPs are of the form $\mathit{M = (S, A, app, Pr, R)}$ where $\mathit{S}$ and $\mathit{A}$ are a finite set of states and actions, $\mathit{app(s)}$ is the set of all actions applicable in $\mathit{s}$. $\mathit{Pr}$ represents the probability of a state transaction $\mathit{(s, a, s')}$ for $\mathit{a}$ and $\mathit{R(s, a, s')}$ is the reward of applying $\mathit{a}$ to $\mathit{s}$ and generating $\mathit{s'}$ The planning problem is defined as a triple $\mathit{P = (M, s\textsubscript{0}, G, \rho)}$ where $\mathit{M = (S, A, app, Pr, R)}$ is an MDP, $\mathit{s\textsubscript{0} \in S}$ is the initial state, $\mathit{G \subseteq S}$ is the set of goal states and $\mathit{0 < \rho < 1}$ is a probability threshold \cite{teichteil2010incremental}.

 The planners from IPPC-04 were evaluated on how many trials reached the goal per problem and the average number of actions selected when successful. These planning problems were however relatively easy to solve as they did not account for \emph{natural dynamics} and/or large branching factors that occur in more realistic systems \cite{kolobov2012glutton}. FF-Replan \cite{FFReplan}, as specified in Section \ref{sec:planners} kindly remembered the committee of the IPPC-04 that the specified planning problems that contained no dead-ends were fairly easy to solve using relatively simple tools to handle with uncertainty.

\subsubsection{MDPs with Finite Horizon. }IPPC-11 decided to respond by introducing more realistic MDPs and their evaluation criteria by using MDPs with finite horizons. The planning problems are defined as tuples of the form \\ $\mathit{M\textsubscript{s0}(H) = \langle \langle S, A, T, R, s\textsubscript{0} \rangle , H \rangle }$ where $\mathit{S}$ s a finite set of states, $\mathit{A}$ a finite set of actions, $\mathit{T}$ is the transition function, similar to $\mathit{Pr}$ in the goal oriented MDP. $\mathit{R}$ represents the transitioning reward, $\mathit{s\textsubscript{0}}$ and $\mathit{H}$ is the number of decision epochs after which the process stops. Solving $\mathit{M(H)}$ means finding a policy for executing recommended state from the initial state over $\mathit{H}$ decision epochs \cite{kolobov2012gourmand}.

Translating finite horizon MDPs to goal-oriented domains results ends up in all state $H$ steps away from $s\textsubscript{0}$ are goal states. The evaluation criteria in finite horizon MDPs is the \emph{expected reward} and not the probability of the policy reaching the goal states.

\subsection{Planning algorithms}
\subsubsection{Replanning}
A replanner is a planner that takes a problem as input, plans a state-action policy, executes it online and whenever it encounters a state that was not in its state-action policy, it takes the current state as starting state and plans again until it has reached a goal or dead-end state. For our paper have chosen to evaluate FF-Replan \cite{FFReplan}, the winner of IPPC-04 and unenrolled winner of IPPC-06 as an inspiring, fairly simple planner that in our opinion is capable of employment in real-world domains.

\subsubsection{Probabilistic planning}
Probabilistic planners are used to solve probabilistic planning problems and should be able to account for uncertainty as opposed to classical planners. For our evaluation we have chosen to compare FF-Replan to a state-of-the-art probabilistic planner: \textsc{Gourmand}. \textsc{Gourmand} is the online successor of the \textsc{Glutton} plan, a Labelled Real-Time Dynamic Programming (LRTDP) based planner and the runner up of IPPC-11.

\subsection{Probabilistic Interestingness}
After IPPC-06, a baseline test for \emph{probabilistically interesting} planning domains was proposed by Little and Thi\'ebauw \cite{little2007probvsreplan}.  Problems that fail this test were deemed to be classical planning problems in disguise, favouring the use of replanners. A probabilistic planning problem is considered to be probabilistically interesting if and only if it has all the following structural properties \cite{little2007probvsreplan}:
% reference little & thiebauw again. Maybe also shortly define what all o, n, act etc. are.

\begin{itemize}
	\item there are multiple goal trajectories;
	\item there is at least on pair of distinct goal trajectories, $\tau$ and $\tau$ ', that share a common sequence of outcomes for the first n-1 outcomes, and where $\tau$\textsubscript{n} and $\tau$'\textsubscript{n} are distinct outcomes of the same action; and
	\item there exist two distinct goal trajectories $\tau$ and $\tau$' and outcome o $\in$ $\tau$ and o' $\in$ $\tau$' of two distinct actions a = act(o) and a'= act(o') such that executing a strictly decreases the maximum probability of reaching a state where a' can be executed.
\end{itemize}

%---------------------------------------------------------------------------------------------------------------------------------

\section{Domains}
\label{sec:domains}

To identify unique properties of the domain and identify properties of the
domain that work particularly well for a certain type of planner, we describe
other domains used in the IPPC and compare that with the UAV domain. Since
Little and Thi\'ebauw's baseline test for \emph{probabilistic interestingness}
of domains, domains of the IPPC competition have shifted to those
probabilistically interesting domains, which are difficult for replanners.
Often domains that don't pass this test are deemed to be classical planners in
disguise, favoring the use of replanners.

\subsection{IPPC Domains}

\subsubsection{Blocksworld}

A simple domain used in the competition of 2004 is the well known blocksworld
domain. The uncertainty lies in that the gripper, used to move blocks, has a
probability that it drops the blocks to the table. Another difference is that
the blocks are colored, and the goal is to build a tower of certain colored
blocks, rather than a tower with specific blocks. If for example there are
four blue blocks, and the goal is a tower of three blue blocks, it does not
matter which blue blocks are used.

\subsubsection{Navigation}

The navigation domain represents a grid in which a robot must reach a goal.
Every cell offers the robot a different chance of disappearing. The robot needs
to choose a path which gets it to the goal most reliably within the finite
horizon time \cite{sanner2011navigation}.

\subsubsection{Traffic}

This domain represents a simple binary version of the cell transition model
(CTM) for modelling traffic based on the original CTM Tech Report
\cite{sanner2011traffic}. Because of the binary variables and no intermediate
variable restrictions, this model is simplified and ignores traffic aspects
such as turns and turn probabilities. This model also uses concurrent actions
and the total actions has an upper bound of $2^{\#~of~intersections}$.

\subsubsection{Elevators}

The IPPC-06 has an elevators domain, however we describe the latest IPPC-11
elevators domain. In this domain there are a number of elevators delivering
passengers to either the top or bottom floor. Potential passengers arrive at
any floor at a certain rate (depending on which floor). The elevators can only
move in the direction it is indicating, so passengers know whether to board or
not. The goal of the elevators is to transport as many as possible passengers
to their destinations.

\subsubsection{SysAdmin}

The SysAdmin domain describes a network of connected servers. Each server has
a probability that it shuts down, and if is not running, it has a probability
to reboot. The probability that a server shuts down depends on the number of
running connected servers.

\subsection{Domain Evaluation}

The Blocksworld domain is ideal for
replanning~\cite{FFReplan, little2007probvsreplan}. As the blocks fall on the
table, which then can be picked up again, it has no dead ends. This makes it
probabilistically uninteresting. For this domain FF-Repan performs really well.

The IPPC-11 domains, Navigation, Traffic, Elevators and SysAdmin are finite-
horizon domains. Navigation, however, is actually a goal oriented domain, if
the goal is reached in the horizon. The method how this is implemented is that
the goal state is defined as an absorbing state, and there is no reward or cost
associated to staying in the goal state, until the horizon is reached.

\textsc{Gourmand} vastly outperformed IPPC-11 UCT-based probabilistic track
winner PROST by far on goal-oriented domains Navigation and Crossing Traffic
\cite{kolobov2012gourmand}. PROST however mildly outperforms \textsc{Gourmand}
on the Game of Life, SysAdmin and the Traffic domain as UCT does not need to
perform expensive Bellman Backups on these domains that include large branching
factors.

The reason \textsc{Gourmand} performs quite well on the navigation domain is
because it uses a dynamic lookahead.

The domains are constructed such that it will challenge planners to perform
better. This causes that there are issues with the domains when comparing them
with the real world.

In the navigation problem there is a chance that the robot
disappears after entering a cell. The disappearing action is a dead end.
Disappearing here is a metaphor for the robot breaking down, or that it
disappears from the radar with the base station. In many of such cases it's a
big problem, but can be repaired with a high cost. In the navigation domain a
step to another cell has a cost of 1. If instead of a dead end the cost is $n$
(e.g. 5) instead of 1, which represents the cost of repairing the robot, dead
ends are removed. Without dead ends the problem becomes probabilistic
uninteresting, which makes it suitable for replanning planners.

The traffic domain describes a crossroad with traffic lights, modelled as
cells. Either the north to south road has a green light, the west to east road
or all lights are red. The model uses a finite state machine that handles the
lights so that the lights are never green at the same time. This would be the
only dangerous situation on the crossroad. The difficulty in this domain is
the number of states, which grows exponentially with the number of cells.
Besides that it is an uninteresting probabilistic problem because it does
not satisfy the probabilistic interesting test.

The UAV domain has a resemblance of the Navigation domain. Both are goal
oriented and have to navigate something (either the robot itself, or a data
package) from an initial state to a goal. Also after each step a dead end can
be reached, the robot disappears or the data package is not delivered.
The probability this happens is static in the navigation domain, but dependent
on the neighbors in the UAV domain. This variable probability has a resemblance
of the SysAdmin domain, where the probability that a server shuts down is also
dependent on the neighbors.

%---------------------------------------------------------------------------------------------------------------------------------

\section{Planners}
\label{sec:planners}

FF-Replan \cite{FFReplan}, a replanner based on determinization techniques and a classical planner, surprisingly outperformed other participating MDP solvers in the IPPC-04 and by doing so changed the course of the planning competition entirely. By relying on search and removing all probabilistic information from the problem, FF-replan functions with enormous speed and efficiency. This method has proven to outperform MDP solvers, the state-of-the-art probabilistic planners for uncertainty of that time. Since then, Little et al. has argued for \emph{probabilistically interesting} domains. By introducing finite horizon MDPs in IPPC-11, state of the art planners have been adapted to cope with large branching factors to approximate infinite planning problems. Because these ``planning problems'' can often be solved using a smaller horizon $h < H$ as introduced in Section \ref{sec:background}, they represent optimization problems rather than planning problems, favoring current state of the art planners over replanners. This section introduces the shift of the planners towards optimization  by discussing FF-replan and a current state of the art probabilistic planner \textsc{Gourmand} and its predecessors.

\subsection{FF-replan}
\label{sec:ffreplan}
 FF-Replan generates a deterministic planning problem by determinizing the input domain at the start of each planning iteration. This classical deterministic planning problem allows for the use of the classical FF planner \cite{Hoffmann01theff}. The FF planner then returns an ordered state-action plan for the deterministic problem of the current state. This plan is executed until an unexpected state is encountered in the following manner:

\begin{enumerate}
	\item Start planning at $s_0$
	\item Determinizes the input domain
	\item Use FF on the classical planning problem
	\item Executes ordered state-action plan until
	\item Unexpected state is encountered or
	\begin{itemize}
		\item Start planning at $s_t$ at time t
		\item Repeat step 2 - 6
	\end{itemize}
	\item Goal state or dead-end is reached
\end{enumerate}

FF-Replan has some obvious shortcomings. It removes all probabilistic information from the problem when determinizing the domain. The success of the planner is therefore fully reliable on the chosen determinization heuristic. As FF-replan uses no roll-out policy or future simulation, it has no way of avoiding \emph{avoidable} dead-ends. This explains why FF-replan performs poorly in domains containing dead-ends.

For a planner to be successful in finite-horizon MDPs it needs to be scalable and sensitive to an expected reward maximization, the evaluation criteria of these domains. To cope with these domains, trial based methods such as \emph{Labelled Real-Time Dynamic Programmers} (LRTDPs) have been introduced and optimized to find an optimal policy in a dynamic programming-based way.

%----------------------------------------------------------------------------------------

\subsection{LRTDP algorithms}

The LRTDP algorithm tries to reach the goal using a series of trials starting from $s_0$ and is specifically convenient for only memorizing states reachable from $s_0$ and therefore uses and needs less memory space in the following manner:

\begin{enumerate}
	\item Start trial at $s_0$
	\item Take action based on current value function $V$
	\item Perform Bellman backup
	\item Sample the outcome and update $V$
	\item Transition to the next state
	\item Repeat step 2 - 5 until a goal, dead end or converged state is reached
	\item Check if all states have converged, otherwise start a new trial (step 2 - 6)
\end{enumerate}

Adapting LRTDP to a finite horizon MDP $M(H)$ solver \emph{LRTDP-FH} requires for each trial to start at $s_0$ or $M(s_0, H)$ and run for at most $H$ time steps. All states $H$ steps away from $s_0$ would translate to goal states in a goal-oriented MDP. \\

The \emph{Labeled Reverse Real-Time Dynamic Programmer} (LR$^2$TDP) uses \emph{reverse iterative deepening} and LRTDP-FH to incrementally solve a sequence of MDPs $M(1)$, $M(2)$, \ldots , $M(H)$ by relying on the fact that a successful policy can be produced by solving a MDP-FH $M(h)$ for a smaller horizon $h < H$.

\begin{enumerate}
	\item Run LRTDP-FH to solve $M(1)$ or $M(s_0, 1)$
	\item Run LRTDP-FH to solve $M(h)$ or $M(s_0, h)$ where $h < H$ by backchaining from the horizon $H$
	\item Repeat step 2 until $H$ is reached or time has run out
\end{enumerate}
By iterating over the distance from the horizon the state values computed when solving $M(1)$, $M(2)$, \ldots, $M(h-1)$ can be reused. Solving $M(h)$ thus only takes an incremental effort over the solution of $M(h-1)$. If stopped prematurely  (LR$^2$TDP) is likely to provide a better policy as for many MDPs the optimal policy $\tau^*$ is optimal or near-optimal for $M(H)$. By itself,  (LR$^2$TDP) would however not perform well in MDPs with large branching factors as computing the Bellman backups would become too expensive carried out over a large fraction of the state space.

\subsection{\textsc{Glutton}}
The \textsc{Glutton} planner enhances LR$^2$TDP to cope with large branching factors by using several optimization techniques. It employs LR$^2$TDP in an off-line fashion by initializing the algorithm with the \emph{Max-Reward} heuristic that places an upper bound on the optimal value function $V^*$. It optimizes for large branching factors by performing the transition function in a subset of transition states $s'$. It distinguished between \emph{exogenous} effects caused by natural dynamics and \emph{pure} effects caused by the action itself and a mixed of both. It also utilized the saved memory space caused by mapping only the converged states in the hash-table by using the free space as cache for storing successor states computed using the transition function.
Since \textsc{Glutton} subsamples the transition function, it may not know a good action to perform in states it missed due to subsampling. To compensate for these states, \textsc{Glutton} has to resort to a \emph{default policy}. \todo{Wel of niet de default policy uitleggen}.
%---------------------------------------------------------------------

When however dealing with large branching factors, offline planning is often a waste of time and memory space since there is a vast amount of states the planner may never reach in a lifetime. Such problems are far better solved using online planners that find an optimal policy $\tau^*$ or an approximation of this policy for the \emph{current} state.

\subsection{\textsc{Gourmand}}
The \textsc{Gourmand} is an online adaptation of the \textsc{Glutton} planner and ensures a good \emph{anytime performance} using a lookahead value $L$ to approximate a near-optimal policy for $M(H)$ where $L < H$ and is dynamically chosen. At every iteration over the distance to the horizon $H$, it divides the total remaining time $\overline{T}$ equally amongst the remaining iterations and virtually allocates time: $T_t = \frac{\overline{T}}{H - t}$. During the first epoch, or iteration it estimates how long solving a state takes for different lookahead values. Using this estimate and a virtual time allocation \textsc{Gourmand} determines the largest lookahead value $L_t$ for which it can almost certainly solve the current epoch  and whether it could solve for a larger lookahead $\widehat{L}_t$ by taking away time from the next epoch, still guaranteeing that it can solve both epochs. If so, it solves the current epoch for $\widehat{L}_t$ and otherwise $L_t$. This way \textsc{Gourmand} never has to resort to default policies and converges relatively fast to a good policy in an online fashion. The \textsc{Gourmand} planner has proven to outperform the current state of the art planners utilizing its robust any-time performance \cite{kolobov2012gourmand} in the following manner:

\begin{enumerate}
	\item Start \textsc{Gourmand} at $s_0$
	\item Initialization of first epoch
	\begin{enumerate}
		\item Virtually allocate time for  $T_t = \frac{\overline{T}}{H}$
		\item Run LR$^2$TDP($M_s(H), T_t$)
		\item Adjust running average $\overline{T}$
	\end{enumerate}
	\item Virtually allocate time  $T_t = \frac{\overline{T}}{H - t}$
	\item Check if lookahead can be increased $\widehat{L}_t = L_t + 1$ else $\widehat{L}_t = L_t $
	\item Run LR$^2$TDP($M_s(\widehat{L}), \widehat{T}_t$)
	\item Adjust running average $\overline{T}$
	\item Determine action based on (near) optimal $\tau^*$
	\item Transition to next state
	\item Repeat steps 3 - 8 until time has run out
\end{enumerate}

\todo{write some kind of conclusion}

%---------------------------------------------------------------------

\section{Future Experiment}
\label{sec:experiment}

Experiments can empirically evaluate the performance of the adjusted domains.

\subsection{Setup}

The planning problems are taken from the IPPC-11 suite and are written in RDDL,
which are publicity available. A domain consists of the domain description and
instances of a domain problem. The domain description is read by RDDLSim, which
is a server a planner can communicate with, and simulates the real world after
the planner sends an action.

\subsection{FF-Replan with Finite Horizon}

\textsc{Gourmand} already works with RDDLSim and MDPs with finite horizon. The original
FF-Replan is a Goal-Oriented MDP planner. For a Finite-Horizon each state at
the horizon is a goal. To compare \textsc{Gourmand}, that works with MDP-FH, with
FF-Replan, FF-Replan needs a slight update. It should not only execute FF for
one goal, but for all goals at the horizon. It could then start executing the
plan with the highest reward. Depending on the determinization method,
\emph{single-outcomes} or \emph{all-outcomes}, not all goals might be
reachable. Of course the performance decreases with an increase of goal states
at the horizon. A hash-table cache, as FF-Replan already uses can still be used
to store plans generated by FF from a state to goal states, however not only
the plan should be stored, but also the reward for this plan should be stored
to pick the plan with the highest expected reward.

Using the framework of \textsc{Glutton} and \textsc{Gourmand}, and the FF heuristic found in mGPT
\cite{bonet2011mgpt}, the updated FF-Replan can be implemented. The first step
of \textsc{Gourmand} is parsing the problem and creating a structure of the problem.
Instead of executing the \textsc{Gourmand} algorithm, this structure can be used to
determinize the problem by setting all probabilities of each action to 1
(\emph{all-outcome}) or all but one to 0 (\emph{single-outcome}). Using the
FF heuristic it can find plans for each goal and calculate the expected reward.
Then it can reuse the \textsc{Gourmand} framework to execute the plan by sending the
action to the RDDLSim server. Like FF-Replan if the server returns with a state
that is not in the plan the algorithm will replan. If a goal is reached it will
output the achieved reward.

\subsection{Hypothesis}

%---------------------------------------------------------------------

\section{Related Work}

In sections above we have described the GOURMAND planner and FF-Replan as two
probabilistic planners of the IPPC competition.
RFF~\cite{teichteil2010incremental} is an offline planner that generates a
policy: for each reachable state it determines a plan to the goal. Like
FF-Replan it uses determinization and FF\. FF returns an initial plan from the
initial state to a goal state. For each state in the plan the successor states
are added to the structure, after which Monte-Carlo sampling is used. States
where the probability that they're reached is too low (below a threshold) are
removed. The successor states are used as initial states in the next step. If
FF fails to find a plan to the goal from this new state, the state is removed
from the structure too. Results showed that RFF outperformed other competitors.
A higher threshold means that RFF will generate more policies that are prone to
replanning.

\subsubsection{PROST ?}

\subsection{Probabilistic planning vs Replanning}
Inspired by the FF-Replan entrance in the IPPC competition, this paper described a probabilistic interestingness quality of domains. It then compares probabilistic planners with replanners in the different domains to draw conclusions. Classical planning with replanning upon failure when there is so little or so much uncertainty about the world that it is not worth modelling and reasoning about or when goals dynamically change or when the problem is too large to solve at once. Probabilistic planners are more appropriate when avoiding irreparable or costly failures. Probabilistic planning computes a plan that can handle all foreseeable contingencies and is preferable for quick reactions in closed control loops.

The following structural properties are identified for comparing probabilistic planning with replanning:
\begin{itemize}
	\item `dead' end states
	\item the degree to which the probability of reaching a dead end state can be reduced through choice of actions
	\item number of distinct trajectories
	\item presence of mutual exclusion
\end{itemize}

If there are no dead ends, a replanner will always succeed. It is impossible to improve on the unavoidable dead ends (positive probability of reaching when executing which cannot be reduced without reducing the probability of reaching the goal).

% repetition of what's already stated earlier

A probabilistic planning problem is considered to be probabilistically interesting if and only if it has all the following structural properties:

\begin{itemize}
	\item there are multiple goal trajectories;
	\item there is at least on pair of distinct goal trajectories, $\tau$ and $\tau$ ', that share a common sequence of outcomes for the first n-1 outcomes, and where $\tau$\textsubscript{n} and $\tau$'\textsubscript{n} are distinct outcomes of the same action; and
	\item there exist two distinct goal trajectories $\tau$ and $\tau$' and outcome o $\in$ $\tau$ and o' $\in$ $\tau$' of two distinct actions a = act(o) and a'= act(o') such that executing a strictly decreases the maximum probability of reaching a state where a' can be executed.
\end{itemize}

Problems that fail to be \emph{probabilistically interesting} are actually classical planning problems in disguise and for those replanners are more appropriate. A replanner can also perform optimally for probabilistically interesting planning problems whenever the `most promising' goal trajectory is the correct thing to do. They create several domains to explore the issue of replanning vs\@. probabilistic planning.

Following existing domains were said to be probabilistically interesting:
\begin{itemize}
	\item Drive
	\item Exploding blocksworld
	\item Pitchcatch
	\item Schedule
	\item Tireworld
\end{itemize}

They conclude by stating that the negative perception of probabilistic planning by the ICAPS community was inaccurate. They state that a synthesis of planning and replanning techniques could make a larger number of probabilistic planning problems practicably solvable than is currently possible.

%---------------------------------------------------------------------

\section{Conclusion}

\begin{itemize}
	\item Conclude the work done in the paper
	\item Hint to some future work developments
\end{itemize}

\bibliographystyle{plain}
\bibliography{paper}

\end{document}